### 2023-01-14 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@2bbe8eebead4e02de00580ad1b3ebabcb23e734b 🚀

        {\bfseries Abstract:}\,Ecosystems are composed of networks of
interacting species. These interactions allow communities of species to
persist through time through both neutral and adaptive processes.
Despite their importance, a robust understanding of (and ability to
predict and forecast) interactions among species remains elusive. This
knowledge-gap is largely driven by a shortfall of data---although
species occurrence data has rapidly increased in the last decade,
species interaction data has not kept pace, largely due to the intrinsic
difficulty and effort required to sample interactions. This means there
are many interactions between species that occur in nature, but we do
not think this interaction occurs because we have no record of it. These
so-called ``false-negatives'' bias data and hinder inference about the
structure and dynamics of interaction networks. Here, we demonstrate the
realized rate of false-negatives in data can be quite high, even in
thoroughly sampled systems, due to the intrinsic variation in abundances
across species in a community. We illustrate how a null model of
occurrence detection can be used to estimate the false-negative rate in
a given dataset. We also show how to directly incorporate uncertainty
due to observation error into model-based predictions of interaction
probabilities between species. One hypothesis is that interactions
between ``rare'' species are themselves rare because these species are
less likely to encounter one-another than species of higher relative
abundance, and that this can (in part) explain the common pattern of
nestedness in bipartite interaction networks. However, we demonstrate
that across several datasets of spatial or temporally replicated
networks, there are positive associations between species co-occurrence
and interactions, which suggests these interactions among ``rare''
species actually exist but simply are not observed. Finally, we assess
how false negatives influence various models of network prediction, and
recommend directly accounting for observation error in predictive
models. We conclude by discussing how the understanding of
false-negatives can inform how we design monitoring schemes for species
interactions.\\%

@@

One interesting result seen in fig.~\ref{fig:addedfnr}(A\&B) is that the
ROC-AUC value does not approach random in the same way the PR-AUC curve
does as we increase the added FNR. The reason for this is that ROC-AUC
is fundamentally not as useful a metric in assessing predictive capacity
as PR-AUC. As we keep adding more false-negatives, the network
eventually becomes a zeros matrix, and these models can still learn to
predict ``no-interaction'' for all possible species pairs, which does
far better than random guessing (ROC-AUC = 0.5) in terms of the false
positive rate (one of the components of ROC-AUC). This highlights a more
broad issue of label class imbalance, meaning there are far more
non-interactions than interactions in data. A full treatment of the
importance of class-balance is outside the scope of this paper, but is
explored in-depth in Poisot (2022).

@@

We also consider how the FNR affects network properties. In
fig.~\ref{fig:addedfnr}(C) we see the mean trophic level across networks
simulated using the niche model (as above), across a spectrum of FNR
values. In addition to the clear dependence on richness, we see that
mean trophic level, despite varying widely between niche model
simulations, tends to be relatively robust to false-negatives and does
not deviate widely from the true value until very large FNRs,
i.e.~\(p_{fn} > 0.7\). This is not entirely unsurprising. Removing links
randomly from a food-web is effectively the inverse problem of the
emergence of a giant component (more than half of the nodes are in a
connected network) in random graphs (see Li \emph{et al.} (2021) for a
thorough review). The primary difference being that we are removing
edges, not adding them, and thus we are witnessing the dissolution of a
giant component, rather than the emergence of one. Further applications
of percolation theory (Li \emph{et al.} 2021) to the topology of sampled
ecological networks could improve our understanding of how
false-negatives impact the inferences about the structure and dynamics
on these networks.

@@

>>>>

        {\bfseries Abstract:}\,Ecosystems are composed of networks of
interacting species. These interactions allow communities of species to
persist through time through both neutral and adaptive processes.
Despite their importance, a robust understanding of (and ability to
predict and forecast) interactions among species remains elusive. This
knowledge-gap is largely driven by a shortfall of data---although
species occurrence data has rapidly increased in the last decade,
species interaction data has not kept pace, largely due to the intrinsic
difficulty and effort required to sample interactions. This means there
are many interactions between species that occur in nature, but we do
not know these interactions occur because we have never observed them.
These so-called ``false-negatives'' bias data and hinder inference about
the structure and dynamics of interaction networks. Here, we demonstrate
the realized rate of false-negatives in data can be quite high, even in
thoroughly sampled systems, due to the intrinsic variation in abundances
across species in a community. We illustrate how a null model of
occurrence detection can be used to estimate the false-negative rate in
a given dataset. We also show how to directly incorporate uncertainty
due to observation error into model-based predictions of interaction
probabilities between species. One hypothesis is that interactions
between ``rare'' species are themselves rare because these species are
less likely to encounter one-another than species of higher relative
abundance, and that this can (in part) explain the common pattern of
nestedness in bipartite interaction networks. However, we demonstrate
that across several datasets of spatial or temporally replicated
networks, there are positive associations between species co-occurrence
and interactions, which suggests these interactions among ``rare''
species actually exist but simply are not observed. Finally, we assess
how false negatives influence various models of network prediction, and
recommend directly accounting for observation error in predictive
models. We conclude by discussing how the understanding of
false-negatives can inform how we design monitoring schemes for species
interactions.\\%

@@

One interesting result seen in fig.~\ref{fig:addedfnr}(A\&B) is that the
ROC-AUC value does not approach random in the same way the PR-AUC curve
does as we increase the added FNR. The reason for this is that ROC-AUC
is fundamentally not as useful a metric in assessing predictive capacity
as PR-AUC. As we keep adding more false-negatives, the network
eventually becomes a zeros matrix, and these models can still learn to
predict ``no-interaction'' for all possible species pairs, which does
far better than random guessing (ROC-AUC = 0.5) in terms of the false
positive rate (one of the components of ROC-AUC). This highlights a more
broad issue of label class imbalance, meaning there are far more
non-interactions than interactions in data. A full treatment of the
importance of class-balance is outside the scope of this paper, but is
explored in-depth in Poisot (2022). Further we see, if anything, gradual
decline in the performance of the model until we reach very high FNR
levels (i.e.~\(p_{fn} > 0.7\)). This is consistent with other recent
work (Gupta \emph{et al.} 2023), although it must be considered that the
empircal data on which these models are trained already are almost
certain to already contain false-negatives.

@@

We also consider how the FNR affects network properties. In
fig.~\ref{fig:addedfnr}(C) we see the mean trophic level across networks
simulated using the niche model (as above), across a spectrum of FNR
values. In addition to the clear dependence on richness, we see that
mean trophic level, despite varying widely between niche model
simulations, tends to be relatively robust to false-negatives and does
not deviate widely from the true value until very large FNRs. This is
not entirely unsurprising. Removing links randomly from a food-web is
effectively the inverse problem of the emergence of a giant component
(more than half of the nodes are in a connected network) in random
graphs (see Li \emph{et al.} (2021) for a thorough review). The primary
difference being that we are removing edges, not adding them, and thus
we are witnessing the dissolution of a giant component, rather than the
emergence of one. Further applications of percolation theory (Li
\emph{et al.} 2021) to the topology of sampled ecological networks could
improve our understanding of how false-negatives impact the inferences
about the structure and dynamics on these networks.

@@

\leavevmode\hypertarget{ref-Gupta2023HowMan}{}%
Gupta, A., Figueroa H., D., O'Gorman, E., Jones, I., Woodward, G. \&
Petchey, O.L. (2023). How many predator guts are required to predict
trophic interactions? \emph{Food Webs}, 34, e00269.
